Concentration Inequalities

### 2015-10-09 causal-bandit/observe_then_best.tex
completes sketch of retret bound proof for bandit case

@@

\begin{itemize}
\item Use $24Nlog(4N/\delta)$ samples to estimate $m$
\item If $\hat m > \frac{N^{3/2}}{\sqrt{T}}$ then stop and do UCB with remaining rounds.
\item Else use causal explore-exploit and let the exploration time $h = T^{2/3}\hat{m}^{1/3}log(TK)^{1/3}$ 
\end{itemize}

Since we don't know $\boldsymbol{q}$, and thus $m$, we can't set the exploration time $h$ in advance based on $m$ as we did for the known $\boldsymbol{q}$ case. Instead we first use $24Nlog(4N/\delta)$ samples to estimate $m$. If $\hat m < \frac{N^{3/2}}{\sqrt{T}}$ we will continue with the causal algorithm and let the total exploration time $h = T^{2/3}\hat{m}^{1/3}log(TK)^{1/3}$.

@@

\eq {
R(T) & =  2\delta E\left[R(T)|\text{ not } \cref{eqn:bounds_hold}\right] + (1-2\delta)E[R(T)| \cref{eqn:bounds_hold}]  \\
& \leq 2\delta T + \left(h + T\sqrt{\frac{\hat m}{h} \log\frac{4N}{\delta}} \right)\\
 &  \leq \frac{4 \hat m^{1/3}T}{h}  + h + T \left( \sqrt{\frac{\hat m}{h} \log \left(\frac{2Nh}{\hat m^{1/3}}\right)}\right), \text { letting } \delta = \frac{2 \hat m^{1/3}}{h} \\
&  \leq 4\left(\frac{T}{log(TK)}\right)^{1/3} + h + T \left( \sqrt{\frac{\hat m}{h} \log \left(TK\right)}\right) \\
&  \leq 4\left(\frac{T}{log(TK)}\right)^{1/3} + 2T^{2/3}(2m)^{1/3}log(TK)^{1/3}  \\
&  \leq 8T^{2/3}m^{1/3}log(TK)^{1/3}  \\
}

Now make sure I have all the cases right - then have a go with an implementation.

>>>>

\begin{lemma}\label{lem:m_est_small_enough}
Let $\delta >0$ and $\hat s_i = \min\set{\hat q_i, 1 - \hat q_i}$ and
define $\hat m = \min\set{2 \leq i \leq N : \hat s_{(i)} \geq 1/i}$.
If
\eq{
h \geq 24m \log\frac{4N}{\delta}\,,
}
then
\eq{
\P{\hat m < \frac{2}{3}m} \leq \delta\,.
}
\end{lemma}

\begin{proof}

To over-estimate $m$, we must identify as balanced some unbalanced arms (ie those with $i < m$. For all arms with $i < m$, by definition, $q_i < \frac{1}{m}$

Now by \cref{lem:conc1} we have with probability at least $1 - \delta$ that: \color{red}(this step not quite right - because I haven't said $q_i = 1/m$ as the worst case is that $q_i$ is infintesimally smaller than $1/m$. With smaller $q_i$ the bounds would be slightly tighter but not made up for by the decrease in the mean - ie we are more likely to over-estimate if $q_i = 1/m$ than if $q_i < 1/m$. Need a better way to say this ... ). Should probably just combine this with the other side into one lemma (even though we don't need this side for simple regret. \color{black}

\eq{
(\forall i) \qquad \left| \hat q_i - q_i\right| 
&\leq \sqrt{\frac{6q_i}{h} \log \frac{2N}{\delta}} \\
&= \sqrt{\frac{6}{mh} \log\frac{2N}{\delta}}\,. 
}

By assumption 
\eq{
h \geq 24m \log \frac{4N}{\delta}\,.
}
Therefore with probability at least $1 - \delta$ we have
\eq{
(\forall i < m) \qquad & \left|\hat q_i - q_i\right| \leq \frac{1}{2m}\,.\\
 & \hat{q}_i < \frac{3}{2}m \\
 \implies & \hat{m} > \frac{2}{3}m
}

\end{proof}

@@

\begin{itemize}
\item Use $24Nlog(4N/\delta)$ samples to estimate $m$
\item If $\hat m > \lambda = \frac{N^{3/2}}{\sqrt{T}}$ then stop and do UCB with remaining rounds.
\item Else use causal explore-exploit and let the exploration time $h = T^{2/3}\hat{m}^{1/3}log(TK)^{1/3}$ 
\end{itemize}

There will be 3 cases.
\begin{enumerate}
\item $m < \lambda/2 \implies \hat{m} < \lambda$ with prob $1-\delta$
\eq {
R_T \sim O(T^{2/3}m^{1/3}) + O(\sqrt{NT})\delta \implies \text { want } \delta \leq \frac{T^{1/6}}{\sqrt{N}}
}
\item $m > 3\lambda/2 \implies \hat{m} > \lambda$ with prob $1-\delta$
\eq {
R_T \sim O(T^{2/3}N^{1/3})\delta + O(\sqrt{NT}) \implies \text { want } \delta \leq \frac{N^{1/6}}{T^{1/6}}
}
\item $\lambda/2 < m < 3\lambda/2$, algorithm could end up doing UCB or Explore-Exploit.
\eq{
R_T \sim  & O(T^{2/3}m^{1/3}) + O(\sqrt{NT}) \\
= &  O(\sqrt{NT}) \text { as } m = O(\frac{N^{3/2}}{\sqrt{T}})
}
\end{enumerate}

So if $\delta = \frac{1}{T^{1/6}\sqrt{N}}$ (or $\delta = \frac{1}{T^{1/3}}$) that would be enough concentration around $m$ to choose the correct algorithm ... Is it enough to choose a reasonable total exploration time $h$? It seems that $\delta = \frac{1}{T^{1/3}}$ works.

\begin{algorithm}[H]
\caption{Bandit Regret Algorithm}\label{alg:bandit}
\begin{algorithmic}[1]
\State {\bf Input:} $T, N$
\State $T_1 = 24Nlog\left(4N/\delta\right)$ TODO figure out what $\delta$ should be 
\For{$t \in 1,T_1$}
\State Do nothing and observe $X_{i,t}$ for $i \in \set{1,\ldots,N}$ and $r_t$
\EndFor
\State Compute $\hat q_i = \frac{2}{h} \sum_{t=1}^{h/2} X_{i,t}$
\State Compute $\hat s_i = \min\set{\hat q_i, 1 - \hat q_i}$
\State Compute $\hat{s'} = sorted(\hat{s}) : \hat{s'}_1 \leq \hat{s'_2} \leq ... \leq \hat{s'}_N$
\State Compute $\hat m = \min\set{2 \leq i \leq N : \hat s'_{i} \geq 1/i}$
\If {$\hat{m} > \frac{N^{3/2}}{\sqrt{T}}$}
\State Switch to the standard UCB algorithm with $K = 2N$ arms.
\Else
\State $h = T^{2/3}\hat{m}^{1/3}log(TK)^{1/3}$ 
\For{$t \in T_1,h/2$}
\State Do nothing and observe $X_{i,t}$ for $i \in \set{1,\ldots,N}$ and $r_t$
\EndFor
\State Compute for all $i \in \set{1,\ldots,N}$ and $j \in \set{0,1}$:
\eq{
\hat \mu_{i,j} = \frac{\sum_{t=1}^{h/2} \ind{X_{i,t} = j} r_t}{\sum_{t=1}^{h/2} \ind{X_{i,t} = j}}\,.
}
\State Recompute $\hat q_i = \frac{2}{h} \sum_{t=1}^{h/2} X_{i,t}$
\State Recompute $\hat s_i = \min\set{\hat q_i, 1 - \hat q_i}$
\State Recompute $\hat{s'} = sorted(\hat{s}) : \hat{s'}_1 \leq \hat{s'_2} \leq ... \leq \hat{s'}_N$
\State Recompute $\hat m = \min\set{2 \leq i \leq N : \hat s'_{i} \geq 1/i}$
\State $i'(i) = $ the index of $\hat s_i$ in $\hat s'$
\State Compute $A$ as the subset of infrequently observed arms $\{(i,j):i'(i) \leq \hat m, j = \ind{\hat q_{i} \leq \frac{1}{2}} \}$ with $|A| = \hat m$

\For{$(i,j) \in A$}
\For{$t \in 1,\ldots,h/2\hat m$}
\State Choose action $X_{i,t} = j$
\State Observe reward $r_t(X_{i,t}=j)$
\EndFor
\State Recompute $\hat \mu_{i,j} = \frac{2\hat m}{h} \sum_{t=1}^{ h/2\hat m} r_t(X_{i,t}=j)$ 
\EndFor
\EndIf
\end{algorithmic}
\end{algorithm}

Since we don't know $\boldsymbol{q}$, and thus $m$, we can't set the exploration time $h$ in advance based on $m$ as we did for the known $\boldsymbol{q}$ case. Instead we first use $24Nlog(4N/\delta)$ samples to estimate $m$. If $\hat m < \frac{N^{3/2}}{\sqrt{T}}$ we will continue with the causal algorithm and let the total exploration time $h = T^{2/3}\hat{m}^{1/3}log(2TK)^{1/3}$.

@@

\eq {
R(T) & =  2\delta E\left[R(T)|\text{ not } \cref{eqn:bounds_hold}\right] + (1-2\delta)E[R(T)| \cref{eqn:bounds_hold}]  \\
& \leq 2\delta T + \left(h + T\sqrt{\frac{\hat m}{h} \log\frac{4N}{\delta}} \right)\\
 &  \leq 2T^{2/3}  + h + T \left( \sqrt{\frac{\hat m}{h} \log \left(4NT^{1/3}\right)}\right), \text { letting } \delta = \frac{1}{T^{1/3}} \\
&  \leq 2T^{2/3} + h + T \left( \sqrt{\frac{\hat m}{h} \log \left(2TK\right)}\right) \\
&  \leq 2T^{2/3}  + 2T^{2/3}(2m)^{1/3}log(2TK)^{1/3}  \\
&  \leq 6T^{2/3}m^{1/3}log(TK)^{1/3}  \\
}

Now lets try an implementation ...
